Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-12-07 Phase_II/draft/GAN.tex
before final

\section{Generative Adversarial Networks}

@@

	\caption{\textcolor{blue}{These figures show the effect of the ratio of episodes of Discriminator training per every episode of Generator training. This hyperparameter is termed Discriminator repetition or (discrep) in the figures. The square root of the cumulative mean of the losses are plotted against the training steps for 3 values of this ratio. Understandably, this ratio has a higher impact on the Discriminator loss than the Generator loss. Equal number of episodes of training produces minimum cumulative loss of the Discriminator. The dip in the Discriminator loss during the initial phases of its training can be interpreted as its early success in detecting ``fake" (or generated) images because of a poorly trained Generator. With gradual training of the Generator, the success rate of the Discriminator decreases and eventually approaches saturation with equal probability of being successful in telling ``fake" from real. The continual decrease and eventual saturation of the Generator loss is a result of its training to generate better images with increasing number of steps.}}

@@

Subsequently, the GAN framework was extended to a conditional model \citep{mirza2014conditional}. This new model, known as ``conditional GAN" or cGAN injects a conditioning variable \(y\) into both networks: the Generator now generates \(G(z \mid y)\), and the Discriminator evaluates \(D(x \mid y)\). The adversarial objective becomes  
\begin{equation}
	\centering
	\begin{aligned}
		V(D, G) &= \mathbb{E}_{x,y \sim p_{\rm data}(x)} \left[ \log D(x|y) \right] \\
		&+ \mathbb{E}_{z,y\sim p_{z}(z)} \left[ \log \left( 1-D(G(z|y) \right) \right]
	\end{aligned}
	\label{eq:conditional_GAN}
\end{equation}
In our work, we specifically use this conditioning mechanism by choosing \(y\) to represent the ground-based intensity-interferometry (II) observation patterns: the Generator is trained to produce stellar surface images that not only look realistic but also conform to the measured II correlations, while the Discriminator judges realism \emph{and} consistency with the II data.

\iffalse 
During each training session of the Discriminator, it is presented with a series of real images (as seen in Fig. (\ref{fig:GANinput}) and generated images taken in equal proportion. During the initial phases of its training, as the generated images are sampled from random distribution, the Discriminator finds it easy to distinguish between the real and the generated images. This produces a sharp initial dip in the Discriminator loss function. On the other hand, the Generator loss function, during the initial phases of its training is high due to the poor quality of its generated images. Through the process of alternating training sessions, the Generator gradually learns to do better: generate images which resemble the real images ever more closely. As seen in fig. (\ref{fig:Plot_discrep_loss}), with further training of both the networks, the Generator loss is gradually minimized and the generated images become gradually indistinguishable from the real images. The Discriminator loss, on the other hand, increases from the minimum of the initial phases of training and gradually  saturates to a value halfway between its maximum and minimum values, as expected in the case of a binary classifier.}
\fi

\cite{isola2017image} further observed that combining the cGAN from Eq.~\eqref{eq:conditional_GAN} with the traditional $L_1(G)$ loss \textcolor{blue}{(also known  as the Mean Absolute Error) improves the results, as the Generator is encouraged to produce outputs closer to the target image in a pixel-wise sense. We adopt this approach in the training of our cGAN model by including $L_1(G)$ along with the hyperparameter $\lambda =100$. The total loss function that is minimized is}:
\begin{equation}
	\centering
	L_{tot} = \arg \min_{G} \max_{D} V(D, G) + \lambda \cdot L_1(G)
	\label{eq:total_gen_loss}
\end{equation}
with the choice $\lambda = 100$ and
\begin{equation}
	\centering
	L_1(G) = \mathbb{E}_{x,y,z} \left[ ||{y - G(x,z)}||_1 \right]
	\label{eq:l1_loss}
\end{equation}
This type of network has demonstrated remarkable robustness across a variety of applications. For example, it can generate colored images from grayscale inputs based on architectural labels, transform images from day to night, and even predict maps from satellite data. A more extensive list of applications is provided in \cite{isola2017image}.

\subsection{Generator}
As discussed above, in a GAN the Generator, a deep neural network in iteself, is responsible for producing synthetic data — in this case, images that resemble those of a fast-rotating star. In this work, the Generator is implemented as a U-Net convolutional network \citep{ronneberger2015u}. \textcolor{blue}{The U-Net consists of a symmetric encoder--decoder structure forming a characteristic ``U'' shape: the left (contracting or down-sampling) path, the right (expanding or up-sampling path) and the connecting (bottle-neck) path. The left down-sampling path repeatedly applies 3$\times$3 convolutions (padded to preserve spatial dimensions) followed by LeakyReLU activations and strided convolution (with stride of 2), by progressively halving the spatial resolution while doubling the channel depth (typically 64 $\to$ 128 $\to$ 256 $\to$ 512 $\to$ 1024). At the bottleneck, high-level features are processed without further spatial reduction. The right up-sampling path mirrors this process using 2$\times$2 transposed convolutions (stride 2) for up-sampling, halving the channel count at each level and using ReLU as the activation function for all its layers except the output layer. Additionally, a dropout layer is introduced at the beginning of the upsampling phase to mitigate overfitting of the Generator model \citep{isola2017image}. Critically, long skip connections concatenate feature maps from each encoder (down-sampling) level to the corresponding decoder (up-sampling level), directly injecting high-resolution details into the reconstruction process. This enables the network to ``remember where everything is'' while the deep bottleneck still provides the large receptive field needed to think globally about image structure and semantics. Besides the strided convolutions, modern variants of U-Nets used in state-of-the-art GANs incorporate residual blocks within resolution levels, and frequently add spectral normalization and self-attention at the bottleneck for improved training stability and long-range dependency modelling. These architectural choices allow our U-Net generator to simultaneously recover sharp, high-frequency details (stellar limb edges, limb-darkening profiles, gravity darkening, and rapid-rotation-induced oblateness) and enforce global coherence (overall disk morphology and physical consistency with the observed interferometric visibilities)---making it ideally suited for high-fidelity sky-image reconstruction of fast-rotating stars from sparse ground-based intensity interferometry observations.}

After generating images, the Generator aims to deceive the Discriminator into classifying the generated images as real. The extent to which the Generator succeeds in this deception is quantified by the GAN loss. When the Discriminator is unable to distinguish between the generated and real images (i.e., when the GAN loss is minimized), the Generator is considered to have reached an optimal state. Conversely, if the generated image fails to fool the Discriminator, the Generator produces a new image for further comparison with the real image. Additionally, the Generator's performance is evaluated using another metric known as the L1 loss, which is defined as the mean absolute error between the pixels of the real image and those of the generated image. Balancing the minimization of both the GAN loss and the L1 loss enables the Generator to produce images that are not only realistic but also faithful to the input data.

\subsection{Discriminator}
The Discriminator is tasked with classifying the images produced by the Generator as either real or fake. It takes a real image from the dataset (often referred to as the target image for the Generator) and provides feedback to guide the Generator toward producing more accurate images. In this work, the PatchGAN model \citep{isola2017image} is employed as the Discriminator. Unlike a traditional global classifier, PatchGAN evaluates individual patches of the image, outputting a grid of predictions rather than a single scalar value. \textcolor{blue}{Each element in the grid corresponds to the ``realness'' of one patch of the image under examination of the Discriminator at a time. The final loss is of the Discriminator is the average over all the patch responses. Evaluating the ``realness'' of the input image in terms of its constituent patches facilitates capture of texture/ style and other high frequency components in the image. As compared to a global discriminator, it also reduces the number of parameters in the network thereby helping reduce computation cost. It also works on images with arbitrary sizes.}

The Discriminator’s architecture begins with an initializer that accepts both the input (generated) images and the corresponding real images. Initially, Salt-and-Pepper noise is added to the input images. PatchGAN then reduces the spatial dimensions of the images to extract localized features, ensuring the model focuses on smaller regions. In this downsampling stage, a leaky version of the Rectified Linear Unit (LeakyReLU) is applied in the convolutional layers, similar to the approach used in the Generator.

Subsequently, zero padding is applied, adding rows and columns of zeros around the images to prevent the loss of spatial information during convolution and to facilitate the extraction of deeper features from the downsampled output. Following this, batch normalization is employed to stabilize learning by normalizing activations, and the Discriminator begins classifying each patch as real or fake. This is followed by additional layers involving LeakyReLU activation, zero padding, and convolution, culminating in a final prediction that is fed to the Generator as feedback.

The effectiveness of the Discriminator is measured by its ability to distinguish between real and fake images, quantified through the Discriminator loss. This loss is composed of two parts: one that measures how accurately the Discriminator identifies real images (by comparing predictions to a target value of 1) and another that assesses how accurately it identifies fake images (by comparing predictions to a target value of 0). Together, these loss components ensure that the Discriminator improves its classification performance, which in turn, challenges the Generator to produce increasingly realistic images.

>>>>

\section{Generative Adversarial Networks}
\begin{figure*}
\centering
\includegraphics[width=0.7\linewidth]{fig/FlowDiagram_TikZ.png}
\caption{A schematic representation of the features of a cGAN model used in this work and the iterative process of its training, validation and testing. The process constitutes four broad stages: (1) choice of an appropriate GAN architecture including both the Discriminator and the Generator (not shown in this figure) (2) preparation of the Training, Validation and Testing datasets and (3) Training and Validation of the Model (4) Testing and Evaluation of the Model. The stages (2), (3) and (4) are depicted in this figure. The datasets are prepared in three broad steps: (i) generating the ``ground truth'' images of fictitious fast-rotators $x$, the sparse II images $y$ used as the ``condition'' images in the Model and the generated images $z$ sampled from a Normal Distribution (ii)merging these images into individual files with $(x|y)$ and/ or $(z|y)$ (as seen in an illustrated sample in Fig.(\ref{fig:GANinput})) and generating the full dataset in this process, and finally (iii) partitioning the full dataset into Training Set, Validation Set and the Testing Set. After the iterative training of the Model and its validation process is complete (``Nash point'' of the Minimax Game is reached), the Model is tested using the Testing Set and evaluated.}
\label{fig:FlowDiagram}
\end{figure*}

@@

	\caption{These figures show the effect of the ratio of episodes of Discriminator training per every episode of Generator training. This hyperparameter is termed Discriminator repetition or (discrep) in the figures. The square root of the cumulative mean of the losses are plotted against the training steps for 3 values of this ratio. Understandably, this ratio has a higher impact on the Discriminator loss than the Generator loss. Equal number of episodes of training produces minimum cumulative loss of the Discriminator. The dip in the Discriminator loss during the initial phases of its training can be interpreted as its early success in detecting ``fake" (or generated) images because of a poorly trained Generator. With gradual training of the Generator, the success rate of the Discriminator decreases and eventually approaches saturation with equal probability of being successful in telling ``fake" from real. The continual decrease and eventual saturation of the Generator loss is a result of its training to generate better images with increasing number of steps.}

@@

Subsequently, the GAN framework was extended to a conditional model \citep{mirza2014conditional}. This new model, known as ``conditional GAN" or cGAN injects a conditioning variable \(y\) into both networks: the Generator now generates \(G(z \mid y)\), and the Discriminator evaluates \(D(x \mid y)\). The adversarial objective becomes  
\begin{equation}
	\centering
	\begin{aligned}
		V(D, G) &= \mathbb{E}_{x,y \sim p_{\rm data}(x)} \left[ \log D(x|y) \right] \\
		&+ \mathbb{E}_{z,y\sim p_{z}(z)} \left[ \log \left( 1-D(G(z|y)|y )\right) \right]
	\end{aligned}
	\label{eq:conditional_GAN}
\end{equation}
The conditional variable $y$ in a cGAN can be various additional information including images, labels or text contextual to ``ground truth'' real data $x$. Among various types of cGANs, Pix2Pix GAN with its image-to-image translation design is suitable for the task at hand.
In our work, we specifically use this conditional variable by choosing \(y\) to represent the ground-based intensity-interferometry (II) observation patterns: the Generator is trained to produce stellar surface images that not only look realistic but also conform to the measured II correlations, while the Discriminator judges realism \emph{and} consistency with the II data.

\cite{isola2017image} further observed that combining the cGAN from Eq.~\eqref{eq:conditional_GAN} with the traditional $L_1(G)$ loss (also known  as the Mean Absolute Error) improves the results, as the Generator is encouraged to produce outputs closer to the target image in a pixel-wise sense. We adopt this approach in the training of our cGAN model by including $L_1(G)$ defined in eq.(\ref{eq:l1_loss})
\begin{equation}
	\centering
	L_1(G) = \mathbb{E}_{x \sim p_{\mathrm{data}, z \sim p_z(z)}} \bigl[ \lVert x - G(z \mid y) \rVert_{1} \bigr].
	\label{eq:l1_loss}
\end{equation}
The total adversarial loss function, along with the $L_1$ loss modulated by a hyperparameter $\lambda$ then becomes
\begin{equation}
	\centering
	L_{tot} = \arg \min_{G} \max_{D} V(D, G) + \lambda \cdot L_1(G).
	\label{eq:total_gen_loss}
\end{equation}

This type of network has demonstrated remarkable robustness across a variety of applications. For example, it can generate colored images from grayscale inputs based on architectural labels, transform images from day to night, and even predict maps from satellite data. A more extensive list of applications is provided in \cite{isola2017image}.

\subsection{Generator}
As discussed above, in a GAN the Generator, a deep neural network in iteself, is responsible for producing synthetic data, in this case, images that resemble those of a fast-rotating star. In this work, the Generator is implemented as a U-Net convolutional network \citep{ronneberger2015u}. The U-Net consists of a symmetric encoder--decoder structure forming a characteristic ``U'' shape along with skip connections: the left (contracting or down-sampling) path, the right (expanding or up-sampling path) and the connecting (bottle-neck) path. The left down-sampling path repeatedly applies 3$\times$3 convolutions (padded to preserve spatial dimensions) followed by LeakyReLU activations and strided convolution (with stride of 2), by progressively halving the spatial resolution while doubling the channel depth (typically 64 $\to$ 128 $\to$ 256 $\to$ 512 $\to$ 1024). At the bottleneck, high-level features are processed without further spatial reduction. The right up-sampling path mirrors this process using 2$\times$2 transposed convolutions (stride 2) for up-sampling, halving the channel count at each level and using ReLU as the activation function for all its layers except the output layer. Additionally, a dropout layer is introduced at the beginning of the upsampling phase to mitigate overfitting of the Generator model \citep{isola2017image}. Critically, long skip connections concatenate feature maps from each encoder (down-sampling) level to the corresponding decoder (up-sampling level), directly injecting high-resolution details into the reconstruction process. This enables the network to ``remember where everything is'' while the deep bottleneck still provides the large receptive field needed to think globally about image structure and semantics. Besides the strided convolutions, modern variants of U-Nets used in state-of-the-art GANs incorporate residual blocks within resolution levels, and frequently add spectral normalization and self-attention at the bottleneck for improved training stability and long-range dependency modelling. These architectural choices allow our U-Net generator to simultaneously recover sharp, high-frequency details (stellar limb edges, limb-darkening profiles, gravity darkening, and rapid-rotation-induced oblateness) and enforce global coherence (overall disk morphology and physical consistency with the observed interferometric visibilities)---making it ideally suited for high-fidelity sky-image reconstruction of fast-rotating stars from sparse ground-based intensity interferometry observations.

During the training, the total Generator Loss function $L_G$ including the $L_1$ loss defined in eq.(\ref{eq:l1_loss}) that is minimized is given by
\begin{equation}
\centering
\begin{aligned}
L_{G} \; &= \; - \, \mathbb{E}_{z \sim p_{z}(z)} \bigl[ \log D(G(z \mid y)\mid y) \bigr] \\
& +\; \lambda \; \mathbb{E}_{x \sim p_{\mathrm{data}, z \sim p_z(z)}} \bigl[ \lVert x - G(z \mid y) \rVert_{1} \bigr].  
\end{aligned}
\label{eq:total_gen_loss}
\end{equation}
Here, 
\begin{itemize}
  \item \(x\) denotes a real data sample (e.g.\ the ``ground-truth'' image; here, the synthetically generated fast rotator image) corresponding to condition \(y\), the simulated II observation data.  
  \item \(z\) is a random latent vector drawn from the prior \(p_z(z)\).  
  \item \(G(z \mid y)\) is the generator output (the reconstructed / synthesized image) given \(z\) and condition \(y\).  
  \item \(D(\cdot \mid y)\) is the discriminator’s estimate (probability) that its input is “real,” given the same condition \(y\).  
  \item \(\lambda\) is a hyperparameter controlling the trade-off between adversarial realism and pixel-wise fidelity (typical values depend on the problem, e.g.\ in pix2pix, \(\lambda = 100\)). 
\end{itemize}

\subsection{Discriminator}
The Discriminator is tasked with classifying the images produced by the Generator as either real or fake. It takes a real image from the dataset (often referred to as the target image for the Generator) and provides feedback to guide the Generator toward producing more accurate images. In this work, the PatchGAN model \citep{isola2017image} is employed as the Discriminator. Unlike a traditional global classifier, PatchGAN evaluates individual patches of the image, outputting a grid of predictions rather than a single scalar value. Each element in the grid corresponds to the ``realness'' of one patch of the image under examination of the Discriminator at a time. The final loss of the Discriminator is the average over all the patch responses. Evaluating the ``realness'' of the input image in terms of its constituent patches facilitates capture of texture/ style and other high frequency components in the image. As compared to a global discriminator, it also reduces the number of parameters in the network thereby helping reduce computation cost. It also works on images with arbitrary sizes.

Prior to the down-sampling of data using PatchGAN, each input image is preprocessed with application of Zero Padding followed by batch normalization. The purpose of Zero Padding is to prevent the loss of spatial information and to facilitate the extraction of deeper features from the down-sampled output. Batch normalization is required to stabilize the learning (loss minimization) process. PatchGAN then reduces the spatial dimensions of the images to extract localized features, ensuring the model focuses on smaller regions. In this down-sampling stage, a leaky version of the Rectified Linear Unit (LeakyReLU) is applied in the convolutional layers, similar to the approach used in the Generator. The probability $D(\cdot|y)$ that the patch of the input image represented by $\cdot$ is ``real'' is evaluated through this process. The loss function $D$ of the full input image is obtained by averaging over all the patch responses.

The Discriminator Loss function $L_D$ that is optimized during the training process is given by 
\begin{equation}
\centering
\begin{aligned}
L_{D} \; &= \; - \, \mathbb{E}_{x \sim p_{\mathrm{data}}(x \mid y)} \bigl[ \log D(x \mid y) \bigr] \\
& -\; \mathbb{E}_{z \sim p_{z}(z)} \bigl[ \log\bigl( 1 - D(G(z \mid y)\mid y)\bigr) \bigr]
\end{aligned}
\end{equation}
where the arguments of the $D$ and $G$ functions are as noted in the text following eq.(\ref{eq:total_gen_loss}) This loss is composed of two parts: one that assesses how accurately it identifies fake images (by comparing predictions to a target value of 0) and the other that measures how accurately the Discriminator identifies real images (by comparing predictions to a target value of 1).

The training procedure of these two components of the cGAN model can be outlined as follows:
\begin{itemize}
\item{The discriminator \(D\) is updated by minimizing \(L_D\), keeping \(G\) fixed (so that \(D\) becomes better at distinguishing real vs generated images).}  
\item{The generator \(G\) is updated by minimizing \(L_G\), keeping \(D\) fixed, thus pushing \(G\) to generate images that are (a) judged “real” by \(D\), and (b) close (in pixel-wise sense) to the ground-truth \(x\).} 
\end{itemize}

### 2025-12-07 Phase_II/draft/II.tex
GAN manuscript reviewed and edited
before final

\section{Intensity Interferometry (II) with IACT arrays}
\begin{figure}
  \includegraphics[width=\linewidth]{fig/baseline.png}
  \caption{The tracks of the baselines due to Earth rotation (described in sec.~\ref{sec:earth}) provided by \textcolor{blue}{the hypothetical observation facility of four telescopes arranged in fig.~\ref{fig:teles}} for one night of observation. \textcolor{blue}{The number of baselines scales as square of number of telescopes in the array thus leading to greater coverage of the observational plane and better image reconstruction prospects.}} 
\label{fig:base}
\end{figure}

@@

\subsection{A Fictitious Fast Rotating Star: Our Test Case}
In our work presented here, we attempt image reconstruction of a fast-rotating star using its simulated Intensity Interferometric observation in a cGAN architecture. \textcolor{blue}{Fast-rotating stars are important test cases for understanding various astrophysical processes, including stellar evolution, internal structure, and dynamical behaviour over time. Fast rotation causes stars to adopt an oblate shape, flattening at the poles and bulging at the equator due to the stronger centrifugal force \citep[e.g.,][]{von1924radiative, 1999A&A...347..185M}. Fig.~\ref{fig:image} shows an image qualitatively representing a fictitious fast-rotating star, with brightness (and, therefore, the effective surface temperature) distributed across its surface. The brightness (effective temperature) is highest at the poles and lowest at the equator, a phenomenon known as gravity darkening \citep{lucy1967gravity}. First direct interferometric detection of stellar photospheric oblateness (of Altair) was pioneered by van Belle et al \cite{vanBelle2001} using the Palomar Testbed Interferometer (PTI) and the Navy Prototype Optical Interferometer (NPOI). Gravity darkening due to fast rotation was first observed through interferometric and spectroscopic data from the CHARA Array for the fast-rotating star Regulus \citep{mcalister2005first}.  As pointed out earlier, these two pieces  of work, all using Michelson Interferometry, make a subset of several others \citep{vanBelle2001, DomicianodeSouza2003, mcalister2005first, DomicianodeSouza2005, Monnier2007, Pedretti2009, Zhao2009, Martinez2021}. The first observation of photospheric oblateness (of $\gamma$ Cassiopeiae or $\gamma$-Cas) using Intensity Interfereometry (II) has been recently carried out at VERITAS observatory and is reported by Archer et al. \cite{Archer-arXiv-2025}. Reportage of such observations of other $\gamma$-Cas like targets and other class of FRS by Cherenkov Telescope arrays, such as  the MAGIC array, are expected by 2026. In addition, observation and measurement of gravity darkening using II is the natural next step and is yet to be reported. In this context, our work of reconstructing the image of FRSs from their II-simulated observations using cGAN is an attempt at solving this inverse problem along with mitigation of loss of phase information in II.}

Intensity Interferometry counts the photons arriving at the telescopes from the stellar object. The correlation of these photon arrivals at the telescopes yields the squared visibility Eq.~\eqref{eq:angular_size_meas}, as explained in subsection~\ref{sec:signal}. The left panel of fig.~\ref{fig:ft} shows the signal from the source shown in Fig.~\ref{fig:image} using II, displayed on linear scales. Point to note here is that this figure represents the signal from the source that would be recorded by an infinite number of baselines provided by an infinite number of telescopes on the interferometric plane. In practice, only a small part of this information is available (as seen in right panel of Fig.~\ref{fig:ft}), because one has a finite number of baselines corresponding to the finite number $N_T$ of telescopes at our disposal and a limited observation schedule. We have simulated the II observation of the fictitious star \textcolor{blue}{by a hypothetical observation facility having an array of four IACTs} (correlated with baselines as seen in Fig.~\ref{fig:teles}) over one night. Using this modest amount of signal from one night's observation, we have trained a cGAN to construct the image of the source.
\begin{figure*}
   \centering
  \includegraphics[width=0.85\linewidth]{fig/ellipse1612.jpg}
  \caption{\textcolor{blue}{Example of the input used for training the cGAN model. \textcolor{blue}{The picture on the left shows the source image, which serves as the ground truth or the real data ($x$, as mentioned in the Flow Diagram (Fig.\ref{fig:FlowDiagram} discussed later} during the training. The picture on  the right represents the simulated Intensity Interferometry (II) observation pattern of the source (on the left) using \textcolor{blue}{a hypothetical observation facility having as array of four IACTs} (see Fig.~\ref{fig:teles}) and the tracks of the six telescope baselines (see Fig.~\ref{fig:base}) generated  by these four IACTs due to Earth's rotation during the observation session. \textcolor{blue}{This pattern referred to as $y$, in the Flow Diagram (Fig.\ref{fig:FlowDiagram} discussed later forms the ``condition" during the training to which the GAN model has to conform. Salt-and-pepper noise is added to this pattern for enhancing the robustness of the cGAN model.} Together, these images form a training pair, where the GAN learns to reconstruct a predicted image (modeling of observed signal) similar to ground truth (left) from the noisy baseline signal (right). The grayscale in both images is normalized to the brightest pixel.}}
  \label{fig:GANinput}
\end{figure*}

>>>>

\section{Intensity Interferometry (II) with IACT arrays}
\begin{figure}
  \includegraphics[width=\linewidth]{fig/baseline.png}
  \caption{The tracks of the baselines due to Earth rotation (described in sec.~\ref{sec:earth}) provided by the hypothetical observation facility of four telescopes arranged in fig.~\ref{fig:teles} for one night of observation. The number of baselines scales as square of number of telescopes in the array thus leading to greater coverage of the observational plane and better image reconstruction prospects.} 
\label{fig:base}
\end{figure}

@@

\subsection{A Fictitious Fast Rotating Star: Our Test Case}
In our work presented here, we attempt image reconstruction of a fast-rotating star using its simulated Intensity Interferometric observation in a cGAN architecture. Fast-rotating stars are important test cases for understanding various astrophysical processes, including stellar evolution, internal structure, and dynamical behaviour over time. Fast rotation causes stars to adopt an oblate shape, flattening at the poles and bulging at the equator due to the stronger centrifugal force \citep[e.g.,][]{von1924radiative, 1999A&A...347..185M}. Fig.~\ref{fig:image} shows an image qualitatively representing a fictitious fast-rotating star, with brightness (and, therefore, the effective surface temperature) distributed across its surface. The brightness (effective temperature) is highest at the poles and lowest at the equator, a phenomenon known as gravity darkening \citep{lucy1967gravity}. First direct interferometric detection of stellar photospheric oblateness (of Altair) was pioneered by van Belle et al \cite{vanBelle2001} using the Palomar Testbed Interferometer (PTI) and the Navy Prototype Optical Interferometer (NPOI). Gravity darkening due to fast rotation was first observed through interferometric and spectroscopic data from the CHARA Array for the fast-rotating star Regulus \citep{mcalister2005first}.  As pointed out earlier, these two pieces  of work, all using Michelson Interferometry, make a subset of several others \citep{vanBelle2001, DomicianodeSouza2003, mcalister2005first, DomicianodeSouza2005, Monnier2007, Pedretti2009, Zhao2009, Martinez2021}. The first observation of photospheric oblateness (of $\gamma$ Cassiopeiae or $\gamma$-Cas) using Intensity Interfereometry (II) has been recently carried out at VERITAS observatory and is reported by Archer et al. \cite{Archer-arXiv-2025}. Reportage of such observations of other $\gamma$-Cas like targets and other class of FRS by Cherenkov Telescope arrays, such as  the MAGIC array, are expected by 2026. In addition, observation and measurement of gravity darkening using II is the natural next step and is yet to be reported. In this context, our work of reconstructing the image of FRSs from their II-simulated observations using cGAN is an attempt at solving this inverse problem along with mitigation of loss of phase information in II.

II counts the photons arriving at the telescopes from the stellar object. The correlation of these photon arrivals at the telescopes yields the squared visibility Eq.~\eqref{eq:angular_size_meas}, as explained in subsection~\ref{sec:signal}. The left panel of fig.~\ref{fig:ft} shows the signal from the source shown in Fig.~\ref{fig:image} using II, displayed on linear scales. Point to note here is that this figure represents the signal from the source that would be recorded by an infinite number of baselines provided by an infinite number of telescopes on the interferometric plane. In practice, only a small part of this information is available (as seen in right panel of Fig.~\ref{fig:ft}), because one has a finite number of baselines corresponding to the finite number $N_T$ of telescopes at our disposal and a limited observation schedule. We have simulated the II observation of the fictitious star by a hypothetical observation facility having an array of four IACTs with their relative positions approximating those at VERITAS (correlated with baselines as seen in Fig.~\ref{fig:teles}) over one night. Using this modest amount of signal from one night's observation, we have trained a cGAN to construct the image of the source.
\begin{figure*}
   \centering
  \includegraphics[width=0.85\linewidth]{fig/ellipse1612.jpg}
  \caption{An illustrative example of the input used for training the cGAN model. The picture on the left shows the source image, which serves as the ground truth or the real data ($x$), as mentioned in the Flow Diagram (Fig.\ref{fig:FlowDiagram} discussed later during the training). The picture on  the right represents the simulated II observation pattern of the source (on the left) using a hypothetical observation facility having as array of four IACTs (see Fig.~\ref{fig:teles}) and the tracks of the six telescope baselines (see Fig.~\ref{fig:base}) generated  by these four IACTs due to Earth's rotation during the observation session. This pattern referred to as $y$, in the Flow Diagram (Fig.\ref{fig:FlowDiagram} discussed later) forms the ``condition" during the training to which the GAN model has to conform. Salt-and-pepper noise is added to this pattern for enhancing the robustness of the cGAN model. Together, these images form a training pair, where the GAN learns to reconstruct a predicted image (modeling of observed signal) similar to ground truth (left) from the noisy baseline signal (right). The grayscale in both images is normalized to the brightest pixel.}
  \label{fig:GANinput}
\end{figure*}
